A quantum parallel parallel Markov chain Monte Carlo

### 2021-11-11 main.tex
parallel mcmc setup juicy

\newcommand{\ttheta}{\boldsymbol{\theta}}
\newcommand{\dd}{\mbox{d}}
\newcommand{\ppsi}{\boldsymbol{\psi}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\I}{\mathbf{I}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\G}{\mathbf{G}}

@@

With the rudiments of quantum minimization in hand, we present our quantum parallel parallel MCMC (qp$^2$MCMC). The general structure of the algorithm is the same as that of other parallel MCMC algorithms: the algorithm generates a Markov chain by iterating between proposing and selecting new Markov chain states.  Unlike classical single-proposal MCMC, parallel MCMC proposes many states and chooses one according to the generalized acceptance probabilities of Equation \eqref{eq:probs}. In general MCMC, evaluation of the target density function $\pi(\cdot)$ is the rate-limiting computational step, becoming particularly onerous in big data scenarios.  While parallel MCMC benefits from improved mixing as a function of MCMC iterations, the increased computational burden of $P$ target evaluations at each step can lead to less favorable comparisons when accounting for elapsed wall-clock time. 
qp$^2$MCMC uses quantum minimization to limit the number of target evaluations within the generalized acceptance step with the help of the Gumbel-max trick, which transforms the discrete sampling task $\theta^{(s+1)} \sim$ Discrete$(\pi_0,\pi_1,\dots,\pi_P)$ into a search problem over discrete set of real numbers.

@@

>>>>

\newcommand{\ttheta}{\boldsymbol{\theta}}
\newcommand{\dd}{\mbox{d}}
\newcommand{\ppsi}{\boldsymbol{\psi}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\I}{\mathbf{I}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\G}{\mathbf{G}}
\newcommand{\ppi}{\boldsymbol{\pi}}

@@

With the rudiments of quantum minimization in hand, we present our quantum parallel parallel MCMC (qp$^2$MCMC). The general structure of the algorithm is the same as that of other parallel MCMC algorithms: the algorithm generates a Markov chain by iterating between proposing and selecting new Markov chain states.  Unlike classical single-proposal MCMC, parallel MCMC proposes many states and chooses one according to the generalized acceptance probabilities of Equation \eqref{eq:probs}:
\begin{align*}
\pi_j = \frac{\pi(\ttheta_j) \prod_{k \neq j} q(\ttheta_j,\ttheta_k)}{\sum_{i=0}^K \pi(\ttheta_i) \prod_{k \neq i} q(\ttheta_i,\ttheta_k)} \, , \quad j=0,\dots,K \, .
\end{align*}
 In general MCMC, evaluation of the target density function $\pi(\cdot)$ is the rate-limiting computational step, becoming particularly onerous in big data scenarios.  While parallel MCMC benefits from improved mixing as a function of MCMC iterations, the increased computational burden of $P$ target evaluations at each step can lead to less favorable comparisons when accounting for elapsed wall-clock time.

 \emph{Having successfully generated proposals $\ttheta_j,$ $j=1,\dots,P$ and evaluated the corresponding densities $q(\cdot,\cdot)$, we would like to use quantum parallelism and an oracle gate (say, $\U_\pi$) to efficiently compute the $\pi(\ttheta_j)$s but immediately encounter a Catch-22 when we seek to draw a sample $\ttheta^{(s+1)} \sim$ \emph{Discrete}$(\pi_0,\pi_1,\dots,\pi_P)$.  We can use neither a quantum nor a classical circuit to draw the sample! On the one hand, drawing the sample within a quantum circuit would somehow require that all superposed states have access to all the $\pi(\ttheta_j)$s at once to perform the required normalization.  On the other hand, drawing the sample within the classical circuit would require access to all the $\pi(\ttheta_j)$s, but measurement in the computational basis can only return one.}

 In light of this dilemma, we propose to use the Gumbel-max trick to transform the generalized accept-reject step into a discrete optimization procedure.  From there, we use quantum minimization to efficiently sample from $\ppi$. Crucially, we get away with quantum parallel evaluation of the target $\pi(\cdot)$ over all superposed states because the Gumbel-max trick does not rely on a normalization step: each superposed state requires no knowledge of any target evaluation other than its own.

%qp$^2$MCMC uses quantum minimization to limit the number of target evaluations within the generalized acceptance step with the help of the Gumbel-max trick, which transforms the discrete sampling task $\ttheta^{(s+1)} \sim$ Discrete$(\pi_0,\pi_1,\dots,\pi_P)$ into a search problem over a discrete set of real numbers.

@@

We need

\subsection{qp$^2$MCMC}
